---
title: "Reproducible Research"
author: "Maithili"
date: "April 14, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Intial Setup: Download data source zip file, extract and read it
```{r File Setup,results='hold'}
if (!file.exists('repdata'))
  dir.create('repdata')
  if (!file.exists('./repdata/repdata%2Fdata%2Factivity.zip')) {
  filepath <-
  'https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip'
  download.file(filepath, './repdata/repdata%2Fdata%2Factivity.zip')
  }
  if (!file.exists('./repdata/activity.csv'))
  {
  unzip('./repdata/repdata%2Fdata%2Factivity.zip', exdir = './repdata')
  }
  activity <- read.csv('./repdata/activity.csv'
  , header = TRUE)

```
  
Calculate total steps per date
```{r Calculate total Steps}
library("dplyr")
  activity_steps<-select(activity, steps, date)%>%
  group_by(date)%>%
  summarise(daily_steps = sum(steps, na.rm = TRUE)
  )
```

Histogram, mean and median for steps taken daily
```{r plots}
mean(activity_steps$daily_steps,na.rm = TRUE)
median(activity_steps$daily_steps,na.rm = TRUE)
hist(activity_steps$daily_steps,xlab="Daily steps")
```


 Mean and median for steps taken daily
```{r MeanMedian}
mean(activity_steps$daily_steps,na.rm = TRUE)
median(activity_steps$daily_steps,na.rm = TRUE)

```

Plot histogram for daily step count
```{r histogram}
#Plot histogram for daily step count
hist(activity_steps$daily_steps,
     breaks = 20,
     xlab = "Daily steps")
```


Calculate average steps for every interval and plot the timeseries
```{r Timeseries}
#Calculate mean steps for every interval
activity_ts <- select(activity, steps, interval) %>%
  group_by(interval) %>%
  summarise(daily_steps = mean(steps, na.rm = TRUE))

#Plot
plot(activity_ts
     ,type = "l"
     ,xlab = "Interval"
     ,ylab = "Daily Step Count")
```


Get max step count and its interval
```{r Max}
#Get the max step count and its interval
maxsteps <- max(activity_ts$daily_steps)
activity_ts[activity_ts$daily_steps == maxsteps,]
```

Replace missing values with averate step count for each interval
```{r Imputation}
#Get the mean step count for each interval
steps_mean_interval <- select(activity, steps, interval) %>%
  group_by(interval) %>%
  summarise(daily_steps = mean(steps, na.rm = TRUE))

#Impute missng values with mean step count for each interval
activity_imp <- activity
for (i in 1:length(activity_imp$date)) {
  if (is.na(activity_imp$steps[i]))
    activity_imp[i, 1] <-
      steps_mean_interval[steps_mean_interval$interval == activity_imp$interval[i], 2]
}
```


Plot histogram for the imputed dataset
```{r HistImputation}
#Calculate total steps for each date on imputed data
activity_imp_steps <- select(activity_imp, steps, date) %>%
  group_by(date) %>%
  summarise(daily_steps = sum(steps))


#Plot histogram for the new dataset
hist(activity_imp_steps$daily_steps, xlab = "Daily steps")
```

 Mean and median for steps taken daily
```{r MeanMedianImputation}
#Calculate new mean and median
mean(activity_imp_steps$daily_steps, na.rm = TRUE)
median(activity_imp_steps$daily_steps, na.rm = TRUE)
```

Split the imputed dateset into weekday and weekends
```{r Split}
#Add a column identifying the day of the week(Weekend/Weekday) based on the date
activity_imp$day <- ifelse(
  weekdays(as.Date(activity_imp$date)) == "Saturday"
  | weekdays(as.Date(activity_imp$date)) == "Sunday",
  "Weekend",
  "Weekday"
)

#Split the imputed table into two tables with means for weekend and weekday
StepsPerInterval.weekend <- tapply(activity_imp[activity_imp$day == "Weekend" ,]$steps,
         activity_imp[activity_imp$day == "Weekend" ,]$interval,mean)

StepsPerInterval.weekday <-tapply(activity_imp[activity_imp$day == "Weekday" ,]$steps,
         activity_imp[activity_imp$day == "Weekday" ,]$interval,mean)
```

Plot the stepcounts for weekends 
```{r Weekend vs Weekday}
# Define the panel parameters
par(mfrow = c(1, 2))

# Plot step couunt for weekdays 
plot(
  as.numeric(names(StepsPerInterval.weekday)),
  StepsPerInterval.weekday,
  xlab = "Interval",
  ylab = "Step Count",
  type = "l",
  main = "Weekday Pattern"
)

## Plot step count for Weekends 
plot(
  as.numeric(names(StepsPerInterval.weekend)),
  StepsPerInterval.weekend,
  type = "l",
  xlab = "Interval",
  ylab = "Step Count",
  main = "Weekend Pattern"
)
```


